---
title: "Week 3 Homework Doing Data Science"
author: "Brian Waite"
date: "November 22, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 1. GitHub Cloning (20 points): 

```{r}
# cd Documents\GitHub\Testrepo
# mkdir publicdata
# cd publicdata
# git clone https://github.com/caesar0301/awesome-public-datasets
# cd awesome-public-datasets\Datasets
```

# 2. Data Summary (20 points): 

### 2.a: Read in titanic.csv

```{r}
# unzip titanic.csv.zip
setwd("C:/Users/Brian/Documents/Github/Testrepo/publicdata/awesome-public-datasets/Datasets")
df = read.csv("titanic.csv")
```

### 2.b: Output count of females and males aboard the Titanic. Plot the frequency of females and males. 

```{r}
sexDF = summary(df$Sex)
print(sexDF)
barplot(sexDF, ylim = c(0,600), xlab = "Sex", ylab = "Number Onboard", main = "Sex frequency on Titanic")
```



### 2.c: Use one apply function to output the means of Age, Fare, and Survival.

```{r}
lapply(df[,c(2,6,10)], mean, na.rm=TRUE)
```

# 3. Function Building (30 points)

### Prep step: Load the data in sleep01.csv

```{r}
sleepdata = read.csv("http://talklab.psy.gla.ac.uk/L1_labs/lab_1/homework/sleep_data_01.csv")
```

### Build the function for building the report. 

```{r}
###This function accepts a dataframe in the format of sleep_data_01.csv imported into a dataframe
###It returns a report of the sleep data consisting of the median age, RSES mean, RSES std. deviation 
###(all divided by 5) and the range of Duration. 

sleepFunction = function(x) { 
  medianAge = median(x$Age, na.rm = TRUE)
  minDuration = min(x$Duration,na.rm = TRUE)
  maxDuration = max(x$Duration, na.rm = TRUE)
  meanRSES = mean(x$RSES, na.rm = TRUE)
  sdRSES = sd(x$RSES, na.rm = TRUE)
  report = as.data.frame(t(c(medianAge/5,meanRSES/5,sdRSES/5,maxDuration - minDuration)))
  names(report) = c("MedianAge", "SelfEsteem", "SE_SD", "DurationRange")
  round(report, 2) #returns report rounded to 2 decimal places at most
  }
```

### Run the function using the sleep data provided. 

```{r}
sleepFunction(sleepdata)
```

# 4. FiveThirtyEight Data (30 points): 

### a. Install the fivethirtyeight package

```{r}
#commenting the install code to avoid issues in compiling the knit file

#install.packages("fivethirtyeight")

#load
library(fivethirtyeight)
```

### Find the data set name of the 22nd data set

```{r}

#get a list of all sets
setListing = data(package = "fivethirtyeight")
#find the 22nd set variable name
as.name(setListing$results[22,3])
```

### b. Assign the data identified to a new dataframe df

```{r}
#assign the dataframe to the identified value. Couldn't get this to take programmatically with as.name, call, eval, etc... Not sure why. 
df = fivethirtyeight::college_recent_grads
```

### c. Use a more detailed list of the datasets to print out a URL in a comment to the news story. 

```{r}
#find the record
largeSetListing = read.csv("https://goo.gl/OT8iHa") 
articleURL = as.character(largeSetListing[which(largeSetListing$DATAFRAME_NAME == setListing$results[22,3]),12])
```

I am really happy to hear from this news story that non-STEM majors can also find ways to make money! 

```{r}
print(articleURL)
```

### d. Using R commands, give the dimension and column names of the dataframe

Dimensions of dataframe df: <numrows> <numcols> 

```{r}
dim(df)
```

Column names of dataframe df: 

```{r}
names(df)
```

#5 Data Summary (30 points): 

### a. Write an R command to get the column names and count the number of columns. 

Again, the column names are: 
```{r}
names(df)
```

And the number of columns is the length of that object: 
```{r}
length(names(df))
```

### b. Generate a count of each unique major_category in data frame with frequency of each. 

```{r}
major_count = as.data.frame(table(df$major_category))
names(major_count) = c("Major", "Frequency")
```

### c. Make a barplot of major_count

```{r}
par(las=2)
barplot(major_count$Frequency, names.arg = major_count$Major, col = blues9, horiz = TRUE, xlim = c(0,30), xlab = "Num Occurrences", main = "Frequency of Majors")

title(ylab="Major", line=0, cex.lab=1.0)
```

### d. Write the fivethirtyeight data to a csv file. Make sure that it does not have row labels.

```{r}
write.csv(df, "college_recent_grads.csv", row.names = FALSE)
```

# 6. Codebook (30 points) 

### a. Start a new repository on GitHub for your SMU MSDS homework. (Complete) 

### b. Create a README.md file which explains the purpose of the repository. (Complete) 

### c. In one (or more) of the nested directories, post your RMarkdown script, HTML file, and data from 'fivethirtyeight.' (complete) 

### d. The link to this repository meeting the above requirements is available at: 

https://github.com/Ujustwaite/Testrepo/tree/master/homework

## Homework 3 Complete! 